---
title: Neighbourhood Conditions on Boston Housing Value
  
# Use letters for affiliations
author:
  - name: Matthew Shu
    affiliation: a
  - name: Monica Turkovic
    affiliation: a
  - name: Raissa Putrizulfan
    affiliation: a
  - name: Yonatan Oron
    affiliation: a
  - name: Ethan Hirschowitz
    affiliation: a
address:
  - code: a
    address: University of Sydney

    
# Optional: line of arbitrary text with additional information.
# Could be used, for example, to mention the bibliographic info in a post-print.
# If not specified, defaults to "This version was compiled on \today"
#date_subtitle: Published in *Journal of Statistical Software*, 2018


# Place eg a DOI URL or CRAN Package URL here
doi_footer: "Neighbourhood Conditions on Boston Housing Value"

# Abstract
abstract: |
  The aim of this study was to construct a multiple linear regression model on Boston Metropolitan Area housing value, based on the dataset published by Harrison and Rubinfeld (1978), and evaluate the impact of the neighbourhood characteristics. Optimising BIC lead to a parsimonious model of 10 regressors ($BIC=-710$, $CV: R^2 = 0.78$). It was found that neighbourhood factors had a significant impact with the proportion of the population at a low socioeconomic status having a large effect. Limitations and areas of future research are discussed below. 

# Optional: Acknowledgements
acknowledgements: |
  This template package builds upon, and extends, the work of the excellent
  [rticles](https://cran.r-project.org/package=rticles) package, and both packages rely on the
  [PNAS LaTeX](http://www.pnas.org/site/authors/latex.xhtml) macros. Both these sources are
  gratefully acknowledged as this work would not have been possible without them.  Our extensions
  are under the same respective licensing term
  ([GPL-3](https://www.gnu.org/licenses/gpl-3.0.en.html) and
  [LPPL (>= 1.3)](https://www.latex-project.org/lppl/)).


# Paper size for the document, values of letter and a4
papersize: letter

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 10pt

# Optional: Force one-column layout, default is two-column
#one_column: true

# Optional: Enables lineno mode, but only if one_column mode is also true
#lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
#numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true

# Optional: Bibliography 
bibliography: references

# Optional: Enable a 'Draft' watermark on the document
#watermark: true

# Customize footer, eg by referencing the vignette
footer_contents: "DATA2002"

# Produce a pinp document
output: pinp::pinp
pdf_document:
    extra_dependencies: ["float"]

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{YourPackage-vignetteentry}
  %\VignetteKeywords{YourPackage, r, anotherkeyword}
  %\VignettePackage{YourPackage}
  %\VignetteEngine{knitr::rmarkdown}
---


## Introduction 

Hedonic Price Modelling has been a valuable method to estimate market prices based on associated characteristics separate from the market [@HARRISON197881; @GOODMAN1978471] [1,14]. Accurately predicted outcomes are valuable for many industries, including homeowners’ insurance, mortgages, price index construction, mass appraisal, and other econometric studies of housing markets [13]. 
In one study of hedonic house pricing by Harrison and Rubinfeld (1978), the 1970 census data from the Boston Metropolitan Area was used to find willingness to pay for better air quality. In this study, the same dataset will be used to answer two important questions; (i) what is the predictive power of this hedonic housing price model on housing prices (ii) how does the neighbourhood conditions of an area affect housing value.

## Dataset


## Analysis

Prior to further analysis, the assumptions of linearity, independence, homoscedasticity and normality were checked for the full model. Due to poor linearity within pairwise plots, various transformations were considered. The extreme skew of ZN, B and RAD made them good candidates for individual log transformation. However, it was found that the residuals vs. fitted plot linearity was only fulfilled with a further log transformation of the MEDV i.e. a log-log relationship. The log-transformed outcome is displayed in the residuals-fitted plot in Figure \ref{fig:partial_log_full_lm_residuals}, where the approximately levelled line indicates a good fit of linearity. As this dataset was collected from the 1970 U.S Census and various governmental organisations, it is presumed that the design of the experiment produced observations that were independent of each other. However, the plausible spatial dependency between house values within the Boston Metropolitan area is perhaps a limitation in the independence assumption. Thirdly, while a slight fanning-in in the log-transformed residuals-fitted plot is observed, there is an approximately constant spread over the range of fitted values, indicating homoscedasticity. Finally, as in Figure \ref{fig:partial_log_full_lm_qq} observations in the QQ plot appear to deviate in the extremities, however, the Central Limit Theorem can be applied to satisfy the normality assumption due to the large set of observations in the data set.


## Refined Model

To refine the model, a Bayesian Information Criterion (BIC) was used, which appropriately punishes the use of excessive regressors [5]. This was applied to both an exhaustive, backwards and forwards search to assess the efficacy of models of different sizes and explanatory variables. All three searches, minimising BIC, are in agreement that a model size of 10 regressors omitting AGE, INDUS and ZN provides an appropriate balance between in-sample predictive power and the number of regressors.

The residuals-fitted plot indicates the linearity assumption is fulfilled with slight overestimates near the extremities as in Fig  \ref{fig:refined_lm_residuals}. The residuals also appear to be homoscedastic with only slight fanning. The QQ-plot indicates residuals appear mostly normal with a slight deviation near the extremities as in Fig \ref{fig:refined_lm_qq}. Given 506 samples, the test statistic can still be assumed to follow the expected distribution by appealing to the Central Limit Theorem. 

The refined model was subsequently checked for multicollinearity by assessing the Variance Inflation Factor. All variables had a VIF < 10 indicating multicollinearity is not a major issue for the refined model [6].

You can get model coefficients from Main.Rmd in Model Refinement section. Please include an equation for us.

\begin{equation}
  \begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \\
       &=x^3+3x^2y+3xy^3+x^3. 
       \label{eqn:example} 
  \end{aligned}
\end{equation}

## Results

### Cross Validation

A 100-repeated 10-fold cross-validation was then performed on both the full model and refined model to assess the out-of-sample performance between models. Extremely similar R^2 values of approximately 0.78 indicated that the refined model with fewer regressors did not sacrifice much in the way of predictive power. In the interest of parsimony, the refined model was chosen. (Graph of R2/MAE if able to fit)

### Model Interpretation

Cohen’s f^2 score was used to determine the effect size of individual predictors in the refined multiple regression model [7]. Based on standard cut-offs, LSTAT was found to have a large effect size whilst all other regressors had small or negligible effect size [D].

We have decided to focus primarily on four neighbourhood variables in our problem statement. namely LSTAT, CRIM, PTRATIO and CHAS. Past research has concluded that there is a positive relationship between household income and property values [8]. This was greatly reflected in our data, where Cohen’s f^2 analysis found that LSTAT had the greatest effect on housing prices, as a one unit increase in the proportion of “lower status” individuals in the area corresponds to a 2.9% decrease in median house value. It has been well known that crime decreases the overall value of a property. In our model we see that  a one unit increase in crime rate will result in a 1% decrease in median house value. Past research has reported that school accessibility [10] and quality [11] has a positive effect on the value of property surrounding it, where the closer a house is from a school, the higher its price. Our model states that a one unit increase in  PTRATIO corresponds to a 4% decrease in house prices. Interestingly this is in disagreement with past literature by Aliyu (2016), Seo and Simons (2009). Past research has indicated that waterfront properties mostly sell at a higher price [12]. This was reflected in our model where a  one unit increase in house price corresponds with a 10.5% increase in CHAS.

## Discussion

When analysing the dataset, different contextual limitations of the data were encountered. These limitations were mostly related to the irrelevance of Boston data from the 1970’s to modern housing markets. It was noticeable that many of the predictor variables used would have little impact on housing prices today due to shifts in societal perspectives and attitudes. For example, the B variable (proportion of blacks) would be an outdated and irrelevant metric in modern housing markets. Other factors such as inflation and changes in the volatilities of housing markets also meant that adjustments would have to be made to the model.

Another major issue was multicollinearity amongst the independent variables. Since the independent variables could be classified into categories such as neighbourhood, pollution and accessibility, there was concern that variables in the same category may be highly intercorrelated. Replicating the approach of the study from which the dataset was sourced, a VIF multicollinearity check was used to ensure that the independent variables had sufficiently low intercorrelation[1]. 

In the final model, one of the most significant independent variables was the crime rate (CRIM). There was concern that the crime rate was too general a predictor, since it was expected that different types of crimes would have differing impacts on house prices. Specifically, impacts of violent crime versus “white collar” crime. An article on the Australian housing market found that many major Australian cities were experiencing simultaneous increases in housing prices and crime rates [2]. Whilst this does not necessarily mean that the two are correlated, it does indicate that in modern days, society may perhaps place more value on location and accessibility over perceived safety.  A different study on the relationship between levels of crime and housing prices found that rather than actual crime rates, proximity to crime hotspots had a more significant impact on house prices [3]. Results showed that specific types of crimes had different effects on the housing market. It was found that the crime with the most significant impact on house prices was vandalism, indicating that housing markets are potentially most concerned about crimes that affect the physical appearance of a neighbourhood.

Based on the findings, it is possible to make recommendations for future research on similar datasets. It is important to ensure that independent variables are contextually relevant and reflective of the general population, by using recent data from a large sample size. Additionally, it is vital to minimise multicollinearity in the model. This could be done by ensuring that independent variables are not intercorrelated by selecting predictors from different sources and fields. Lastly, variables should be adjusted to ensure that they accurately measure the impact on the dependent variable. For example, future studies on the impacts of different factors on house prices could consider proximity to crime hotspots or levels of vandalism rather than the general crime rate.

## Conclusion

Overall, the regression model was successful in modelling different impacts on housing prices in Boston in the 1970’s. Research found that variables such as crime rate, social status of the population and the ratio of pupils to teachers, all belonging to the general group of “quality of the neighbourhood” had the largest impact on house prices. Whilst the model was able to predict house prices for that specific housing market, due to the obsolescence of the data, it would be impossible to predict modern house prices. 

## Appendix

INSERT DATA DESCRIPTION TABLE


```{r partial_log_full_lm_residuals, fig.width=2.5, fig.height=2.5, cache=TRUE, echo=TRUE, fig.cap="\\label{fig:partial_log_full_lm_residuals}Residuals plot against fitted values of full model after log transformation.", echo=FALSE, message=FALSE, warning=FALSE, fig.pos='!H', out.extra = ''}
library(leaps)
library(tidyverse)
library(ggthemr)
library(caret)
library(car)

ggthemr('flat',
        text_size = 7)

df=read.csv("data.csv", header=TRUE)

transformed_df = df %>%
  mutate(
    log_ZN = log(1+ZN),
    log_B = log(B),
    log_RAD = log(RAD)
  ) %>%
  select(
    -ZN,
    -B,
    -RAD
  )

partial_log_full_lm = lm(log(MEDV) ~ ., transformed_df)


ggplot(partial_log_full_lm, aes(x = .fitted, y = .resid))+ 
  geom_point() +
  geom_hline(yintercept=0, col="red") +
  geom_smooth(se=FALSE) +
  labs(x="Fitted Values", y="Residuals", title= "Residual vs. Fitted")
```

```{r partial_log_full_lm_qq, fig.width=2.5, fig.height=2.5, cache=TRUE, echo=TRUE, fig.cap = "\\label{fig:partial_log_full_lm_qq}QQ plot of full model residuals after log transformation.", echo=FALSE, message=FALSE, warning=FALSE}
ggthemr('flat',
        text_size = 7)
ggplot(partial_log_full_lm, aes(sample = .resid))+
  geom_qq() +
  geom_qq_line(col="red") +
  labs(x="Theoretical Quantiles", y="Standardised Residuals", title="Normal Q-Q Residuals")
```

```{r refined_lm_residuals, fig.width=2.5, fig.height=2.5, cache=TRUE, echo=TRUE, fig.cap = "\\label{fig:refined_lm_residuals}Residuals plot against fitted values of refined model. Slight difference.", echo=FALSE, message=FALSE, warning=FALSE}
ggthemr('flat',
        text_size = 7)
mod_lm = lm(log(MEDV)~.-log_ZN -INDUS -AGE, transformed_df)
ggplot(mod_lm, aes(x = .fitted, y = .resid))+ 
  geom_point() +
  geom_hline(yintercept=0, col="red") +
  geom_smooth(se=FALSE) +
  labs(x="Fitted Values", y="Residuals", title= "Residual vs. Fitted")

```

```{r refined_lm_qq, fig.width=2.5, fig.height=2.5, cache=TRUE, echo=TRUE, fig.cap = "\\label{fig:refined_lm_qq}QQ plot of refined model residuals. Slight difference.", echo=FALSE, message=FALSE, warning=FALSE}
ggthemr('flat',
        text_size = 7)
ggplot(mod_lm, aes(sample = .resid))+
  geom_qq() +
  geom_qq_line(col="red") +
  labs(x="Theoretical Quantiles", y="Standardised Residuals", title="Normal Q-Q Residuals")

```




## Original Template

**Everything underneath is a template that helps direct how to add various objects to the report.**


This *pinp is not PNAS* template started when the introduction to
[Rcpp](http://dirk.eddelbuettel.com/code/rcpp.html) by \cite{PeerJ:Rcpp} 
was converted into this updated
[Rcpp Introduction](https://eddelbuettel.github.io/pinp/Rcpp-introduction.pdf)
vignette.  It is based on the
[pnas_article](https://github.com/rstudio/rticles/tree/master/inst/rmarkdown/templates/pnas_article)
template of the wonderful [rticles](https://cran.r-project.org/package=rticles) package by
\cite{CRAN:rticles}. The conversion from markdown to latex is facilitated by
[rmarkdown](https://cran.r-project.org/package=rmarkdown)
\citep{CRAN:rmarkdown} and [knitr](https://cran.r-project.org/package=knitr)
\citep{CRAN:knitr}. The underlying LaTeX macros are from
[pnas.org](http://www.pnas.org/site/authors/latex.xhtml).

The remainder of the document carries over from the corresponding
[pnas_article](https://github.com/rstudio/rticles/tree/master/inst/rmarkdown/templates/pnas_article)
template document. but has been edited and updated to our use case.  A
few specific tips follow.  In general, for fine-tuning some knowledge
of LaTeX is helpful.



## Author Affiliations 

Per common academic best practice, you can include your department,
institution, and complete address, with the ZIP/postal code, for each
author. Use lower case letters to match authors with institutions, as
shown in the example. Authors with an ORCID ID may supply this
information at submission.

## Document Options

We support several options via the YAML header

- Setting a DOI or URL footer, for example for the CRAN package URL,
  which is placed in the bottom-left footer of the title page and even pages;
- Setting a footer label, for example _YourPackage Vignette_ stating
  your package, which is placed in the bottom-right footer on odd
  pages;
- Setting a free-form author field used on the inside footer;
- Optional _Draft_ watermark to be added to each page;
- Line of custom text in subtitle (`date_subtitle`) suitable to give publication
info of the draft, e.g. journal name in a post-print;
- Document date that appears in the footer can be specified manually using
`document_date`.



## References 

Here we differ from PNAS and suggest natbib. References will appear in
author-year form. Use `\citet{}`, `\citep{}`, etc as usual.

We default to the `jss.bst` style. To switch to a different bibliography
style, please use `biblio-style: style` in the YAML header.


## Inline R Code 

The PNAS sample included a fixed PNG image here, but this document prefers
to show the results and embedding of _R_ code. 

```{r figex, fig.width=3, fig.height=3, cache=TRUE, echo=TRUE, fig.cap="Narrow ggplot2 figure"}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
    geom_point(size=3, aes(colour=factor(cyl))) +
    theme(legend.position="none")
```

Here we use a standard knitr bloc with explicit options for

- figure width and height (`fig.width`, `fig.height`), both set to three inches;
- whether the code is shown (`echo=TRUE`); and
- the caption (`fig.cap`) as shown above.


## Digital Figures 

Markdown, Pandoc and LaTeX support `.eps` and `.pdf` files.

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

The R examples above show how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. 

One (roundabout) way of doing this is to _not_ actually plot a figure, but to
save it in a file as the following segment shows:

```{r densityPlot, echo=TRUE}
library(ggplot2)
p <- ggplot(data = midwest,
            mapping = aes(x = area,
                          fill = state,
                          color = state)) +
    geom_density(alpha = 0.3)
## save to file
suppressMessages(ggsave("densities.pdf", p))
```

This file is then included via standard LaTeX commands.

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.66\textwidth, height=3.5in]{densities} 
  \end{center}
  \caption{Wide ggplot2 figure}\label{fig}
\end{figure*}


## Typeset Code (But Do Not Run It) 

We can also just show code.

```r
xx <- faithful[,"eruptions"]
fit <- density(xx)
plot(fit)
```

This simply used a pandoc bloc started and ended by three backticks,
with `r` as the language choice.  Similarly, _many_ other languages
can be typeset directly simply by relying on pandoc.


## Single column equations 

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures. The `\begin{widetext}...\end{widetext}` environment as shown
in equation \ref{eqn:example} below is deprecated, but \LaTeX commands
`\onecolumn` and `\twocolumn` work fine.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

\begin{equation}
  \begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \\
       &=x^3+3x^2y+3xy^3+x^3. 
       \label{eqn:example} 
  \end{aligned}
\end{equation}


<!-- pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->

